---
title: "Randomly Sampled Set"
output:
  pdf_document: default
  html_document: default
---
Many thanks to Alboukadel Kassambara at DataNovia for providing teachings on how to use majority of code in project
Lessons taken were from links below:
K-means: https://www.datanovia.com/en/courses/partitional-clustering-in-r-the-essentials/
DBSCAN: https://www.datanovia.com/en/lessons/dbscan-density-based-clustering-essentials/
Hierarchical: https://www.datanovia.com/en/lessons/agglomerative-hierarchical-clustering/
Cluster optimisation: https://www.datanovia.com/en/lessons/determining-the-optimal-number-of-clusters-3-must-know-methods/
Cluster validaiton: https://www.datanovia.com/en/lessons/cluster-validation-statistics-must-know-methods/
Cluster Validation cont: https://www.datanovia.com/en/lessons/choosing-the-best-clustering-algorithms/
Heatmaps: https://www.datanovia.com/en/lessons/heatmap-in-r-static-and-interactive-visualization/

```{r setup, include=FALSE}
library(dbscan)
library(tidyr)
library(tidyverse)
library(factoextra)
library(NbClust)
library(cluster)
library(dplyr)
library(gplots)
library(pheatmap)
library(ComplexHeatmap)
library(ggplot2)
library(dplyr)
library(reshape2)
library(Rtsne)
library(labdsv)
library(ggplot2)
library(ggpubr)
library(vegan)
library(clValid)
library(fpc)
library(data.table)
library(FactoMineR)
```

Functions
```{r, echo = FALSE}

cutter <- function(hcluster, k, original_data) {
  clustered_data <- original_data %>% 
    mutate(cluster = cutree(hcluster, k))
  return(clustered_data)
}

meltcor <- function(filtered_cluster) {
    filtered_cluster <- filtered_cluster[, c(1,2,3,4,5,6,7)]
    filtered_cluster <- round(cor(filtered_cluster, use = "everything"),2)
    filtered_cluster <- melt(filtered_cluster)
  return(filtered_cluster)
}
chmap <-  function(x) {
  name <- match.call()
  hmap <- ggplot(x, aes(x=Var1, y=Var2, fill=value)) + geom_tile() + 
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "light grey", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed()+ ggtitle("Heatmap correlation:", name) + 
  geom_text(aes(Var2, Var1, label = value), color = "black", size = 4)
   return(hmap)
}
```

Reading data in & memory size
``` {r, echo = FALSE}
#2M
fulldata <- read.csv('FYP 2M OBS.csv')
fulldata <- data.table(fulldata)

#100K
sample005_3 <- read.csv('FYP 100k OBS S3.csv')
sample005_3 <- as_tibble(sample005_3)
s005_3df <- scale(sample005_3)

#21K
sample001_3 <- read.csv('FYP 21k OBS S3.csv')
sample001_3 <- as_tibble(sample001_3)
s001_3df <- scale(sample001_3)


#Validation sample
set.seed(124)
V_sample <- fulldata[sample(.N, 7500)]
vdf <- scale(V_sample)

```

KMEANS
``` {r, echo = FALSE}
set.seed(124)
gc()
NbClust(vdf, distance = "euclidean", min.nc = 2, method = "kmeans", index = "silhouette") #2
vkm <- kmeans(vdf, 2, iter.max = 25, nstart = 1) 
fviz_cluster(vkm, vdf, geom = "point", ellipse.type = "convex", ggtheme = theme_bw(), main = "K-Means for sample size 7,500")

#21k
km001_3 <- kmeans(s001_3df, 2, iter.max = 25, nstart = 1)
fviz_cluster(km001_3, sample001_3, geom = "point", ellipse.type = "convex", ggtheme = theme_bw(), main = "K-Means for sample size ~21,000")

#100k
km005_3 <- kmeans(s005_3df, 2, iter.max = 25, nstart = 1)
fviz_cluster(km005_3, sample005_3, geom = "point", ellipse.type = "convex", ggtheme = theme_bw(), main = "K-Means for sample size ~100,000")
```

KMEANS Heatmaps
```{r, echo = FALSE}
#21k~
km001_3c <- sample001_3 %>% mutate(cluster= km001_3$cluster) %>% group_by(cluster)
km001_3c <- as_tibble(km001_3c)
km001_3c1 <- km001_3c %>% filter(cluster == 1)
km001_3c2 <- km001_3c %>% filter(cluster == 2)

#printing the heatmaps
ggarrange(chmap(meltcor(km001_3c1)), chmap(meltcor(km001_3c2)), widths = c(1,1), labels = c("Cluster 1", "Cluster 2"))

#100k~
km005_3c <- sample005_3 %>% mutate(cluster= km005_3$cluster) %>% group_by(cluster)
km005_3c <- as_tibble(km005_3c)
km005_3c1 <- km005_3c %>% filter(cluster == 1)
km005_3c2 <- km005_3c %>% filter(cluster == 2)

#printing the heatmaps
ggarrange(chmap(meltcor(km005_3c1)), chmap(meltcor(km005_3c2)), widths = c(1,1), labels = c("Cluster 1", "Cluster 2"))
```

DBSCAN
``` {r, echo = FALSE}
set.seed(124)
#V_sample
kNNdistplot(vdf, k = 14) %>%
  abline(h = 2.1, lty = 2) + title("kNN plot for validation sample")
vdb <- dbscan::dbscan(vdf, eps = 2.1, MinPts = 14)
fviz_cluster(vdb, vdf, geom = "point", main = "V_Sample Data cluster")

#21K
kNNdistplot(s001_3df, k = 14) %>%
  abline(h = 2.3, lty = 2) + title("kNN plot for sample 21k")
db001_3 <- dbscan::dbscan(s001_3df, eps = 2.3, MinPts = 14)
fviz_cluster(db001_3, s001_3df, geom = "point", main = "DBSCAN for sample size ~21,000")

#100K
kNNdistplot(s005_3df, k = 14) %>%
  abline(h = 2.3, lty = 2) + title("kNN plot for sample 100k")
db005_3 <- dbscan::dbscan(s005_3df, eps = 2.3, MinPts = 14)
fviz_cluster(db005_3, s005_3df, geom = "point", main = "Sample005_3 Data cluster")
```

Heatmap setup DBSCAN
``` {r, echo = FALSE}
#validation
vdbc <- V_sample %>% mutate(cluster= vdb$cluster) %>% group_by(cluster)
as_tibble(vdbc)
vdb_0c <- vdbc %>% filter(cluster == 0)
vdb_1c <- vdbc %>% filter(cluster == 1)
vdb_2c <- vdbc %>% filter(cluster == 2)
#21K
s001_3c <- sample001_3 %>% mutate(cluster= db001_3$cluster) %>% group_by(cluster)

s001_3c <- as_tibble(s001_3c)
s001_3c0 <- s001_3c %>% filter(cluster == 0)
s001_3c1 <- s001_3c %>% filter(cluster == 1)
s001_3c2 <- s001_3c %>% filter(cluster == 2)

#100k
s005_3c <- sample005_3 %>% mutate(cluster= db005_3$cluster) %>% group_by(cluster)

s005_3c <- as_tibble(s005_3c)
s005_3c0 <- s005_3c %>% filter(cluster == 0)
s005_3c1 <- s005_3c %>% filter(cluster == 1)
s005_3c2 <- s005_3c %>% filter(cluster == 2)

```

DBSCAN Heatmaps
``` {r, echo = FALSE}
#Validation set
ggarrange(chmap(meltcor(vdb_1c)), chmap(meltcor(vdb_2c)), chmap(meltcor(vdb_0c)), widths = c(1,1), labels = c("Cluster 1", "Cluster 2", "Outlier Cluster"))

#21k
ggarrange(chmap(meltcor(s001_3c1)), chmap(meltcor(s001_3c2)), chmap(meltcor(s001_3c0)), widths = c(1,1), labels = c("Cluster 1", "Cluster 2", "Outlier Cluster"))

#100k
ggarrange(chmap(meltcor(s005_3c1)), chmap(meltcor(s005_3c2)), chmap(meltcor(s005_3c0)), widths = c(1,1), labels = c("Cluster 1", "Cluster 2", "Outlier Cluster"))

```

HIERARCHICAL
``` {r, echo = FALSE}
set.seed(124)
#Validation sample
fviz_nbclust(vdf, hcut, method = "silhouette", k.max = 10) #sil = 3,

Vdist <- dist(vdf, method = "manhattan")

hcv <- hclust(Vdist, method = "ward.D2")

hcv_c <- cutter(hcv, 2, V_sample)
sub_hcv <- cutree(hcv, 2)

fviz_cluster(list(data = vdf, cluster = sub_hcv, main = "Ward clustering validation"))
fviz_nbclust(s001_3df, hcut, method = "silhouette", k.max = 10)

#validation sample heatmap
hcv_c <- as_tibble(hcv_c)
hcv_c1 <- hcv_c %>% filter(cluster == 1) 
hcv_c2 <- hcv_c %>% filter(cluster == 2)

ggarrange(chmap(meltcor(hcv_c1)), chmap(meltcor(hcv_c2)), chmap(meltcor(hcv_c3)),
          labels = c("Cluster 1", "Cluster 2"), widths = c(1,1) )

#21k sample
dis001_3 <- dist(s001_3df, method = "manhattan")

#clustering
hc001_3 <- hclust(dis001_3, method = "ward.D2")

#cutting
hc001_3c <- cutter(hc001_3, 2, sample001_3)
sub001_3 <- cutree(hc001_3, 2)

#plots
fviz_cluster(list(data = s001_3df, cluster = sub001_3), main = "Hierarchical for sample size ~21,000")

#Heatmaps
hc001_3c <- as_tibble(hc001_3c)
hc001_3c1 <- hc001_3c %>% filter(cluster == 1) 
hc001_3c2 <- hc001_3c %>% filter(cluster == 2)

ggarrange(chmap(meltcor(hc001_3c1)), chmap(meltcor(hc001_3c2)), 
          labels = c("Cluster 1", "Cluster 2"), ncol = 2, nrow = 1)
```

ggPlots
``` {r, echo = FALSE}
ggplot(s001_3c1, aes(DateKey, PriceVsMax)) + geom_jitter()
ggplot(s001_3c1, aes(PriceVsMax, DateKey)) + geom_jitter()
s001_3c1<- as.data.table(s001_3c1)
s001_3c1[,.N, by=s001_3c1$PriceVsMax]


ggplot(hc001_3c2, aes(DateKey, PriceVsMax)) + geom_bar(stat = "identity")
ggplot(hc001_3c2, aes(PriceVsMax, Forecast)) + geom_jitter() + geom_smooth()+ ggtitle("Forecast vs PriceVsMax DBSCAN Cluster 2")
ggplot(hc001_3c2, aes(DateKey, PriceVsMax)) + geom_jitter() + geom_smooth(method = "lm", se = FALSE) + ylim(0,1)+ ggtitle("PriceVsMax vs DateKey Hierarchical Cluster 2")
ggplot(hc001_3c2, aes(PriceVsMax, Forecast)) +geom_jitter() + geom_smooth(method = "lm", se = FALSE) + ylim(0,1.5)+ ggtitle("Forecast vs PriceVsMax Hierarchical Cluster 2")

s001_3c2<- as.data.table(s001_3c2)
s001_3c2[,.N, by=s001_3c2$PriceVsMax]
s001_3c2[,.N, by=s001_3c2$Forecast]

```

Validation
``` {r, echo = FALSE}
#Internal validation (Dunn/silhouette)
#Kmeans
kmstats <- cluster.stats(dist(vdf), vkm$cluster)
kmstats1 <- cluster.stats(dist(s001_3df), km001_3$cluster)
kmstats$dunn
sil_vk <- silhouette(vkm$cluster, dist(vdf, "euclidean"))
fviz_silhouette(sil_vk)
sil_21k <- silhouette(km001_3$cluster, dist(s001_3df, "euclidean")) 
fviz_silhouette(sil_21k)

#DBSCAN
dbstats <- cluster.stats(dist(vdf), vdb$cluster)
dbstats$dunn
sil_vd <- silhouette(vdb$cluster, dist(vdf, "euclidean"))
fviz_silhouette(sil_vd)
sil_21kd <- silhouette(db001_3$cluster, dist(s001_3df, "euclidean")) 
fviz_silhouette(sil_21kd)

#Hierarhical
hcstats <- cluster.stats(dist(vdf), hcv_c$cluster)
hcstats$dunn
sil_vh <- silhouette(hcv_c$cluster, dist(vdf, "manhattan"))
fviz_silhouette(sil_vh)
sil_21kh <- silhouette(hc001_3c$cluster, dist(s001_3df, "manhattan")) 
fviz_silhouette(sil_21kh)

#clValid (cluster comparison)
clMethods <- c("kmeans", "hierarchical")
rownames(vdf) <- NULL
intern_kh <- clValid(vdf, nClust = 2:4, clMethods = clMethods, validation = "internal", maxitems = 7500)
stab_kh <- clValid(vdf, nClust = 2:4, clMethods = clMethods, validation = "stability", maxitems = 7500)

optimalScores(intern_kh)
optimalScores(stab_kh)


#Dimension description
s001.desc <- PCA(sample001_3, graph = FALSE)
res.desc <- dimdesc(s001.desc, axes = c(1,2), proba = 0.05)
res.desc$Dim.1
res.desc$Dim.2

```